---
title: "Lab Week 5: Time series and a map"
author: "Bret Folger"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{r}

#load packages in!

library(janitor)
library(lubridate)
library(tidyverse)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)

```


## Monthly US energy consumption (renewables)

```{r}

us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

```


 - Make the description all lowercase
 - only keep observations where the description variable contains "consumption"
 - remove obdservations where the description variable contains "total"
 
 
```{r}

renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total")) # ! denotes "do the opposite"

```

### Convert 'yyyymm' column to a date

```{r}

renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# make a version where I have the month & year in separate columns

renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))

```


### Look at it:

```{r}

renew_gg <- ggplot(data = renew_date, aes(x = month_sep, 
                                          y = value,
                                          group = description)) + # don't need the group here once you add an aesthetic within the geom_line() function
  geom_line(aes(color = description))

renew_gg

```

Update colord with paletteer palettes:

```{r}

renew_gg + 
  scale_color_paletteer_d("vapoRwave::hyperBubble")

```

333 Coerce our renew_parsed to a tsibble, which is a time series enabled data frame:

```{r}

renew_ts <- as_tsibble(renew_parsed, 
                        key = description,
                        index = month_sep)

```

Let's look at our ts data ina  couple different ways:

```{r}

renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value) # breaks up each description by monthly trends
#renew_ts %>% gg_season(value) # if fancier functions don't work, we can make them in ggplot()

ggplot(data = renew_parsed, aes(x = month, 
                                y = value, 
                                group = year)) + 
  geom_line(aes(color = year)) + 
  facet_wrap(~ description, 
             ncol = 1, # only want one column
             scales = "free", # allows each facet_wrapped graph has their own scale
             strip.position = "right") 

# could knit this so that it's bigger

```

### Just look at the hydroelectric energy consumnption:

```{r}

hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year)) + 
  geom_line(aes(color = year))

```

### What if I want the quarterly average consumption for hydro?

```{r}

hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))

```

### Decompose hydro_ts time series data and look at the different components

```{r}

dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot() # gives us the four: original, trend, seasonal, and random. LOOK AT THE SCALE. Residual is less than 10% of total range - so here it looks okay.

```

```{r}

hist(components(dcmp)$remainder) # they look normally distributed! Which is what we want? I guess so

```

Now look at ACF:

```{r}

hydro_ts %>% 
  ACF(value) %>% 
  autoplot() # it looks like there's some seasonality...hooray! Obdervations that are 12 months apart are more closely related than any other time

```

# DANGER DANGER

```{r}

hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value),# trying to find best combination of pdqPDQ of season values
    ETS(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010)) # will only show you forecasred values unless you add the filter part on

# built in increased uncertainty as we get further and further into the future
```

### Make a world map!

```{r}

world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)

```

